NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-30 paper/application.tex
grammar fix

In Fig.~\ref{fig:predictStability}(a)(b)(c), we compare the overall prediction stability (regarding synonymous perturbation) for all correct predictions in the development set (10k examples in total).
%
We observe a drastic difference between the stability for \emph{entailment} predictions compare to the \emph{contradiction} and \emph{neutral} ones.
%
Such a distinction can be partially explained by how \emph{entailment} relationship is defined. The relationship is only valid if the concept in the premise is more specific than the concept in the hypothesis. Therefore, the synonymous perturbation may change the \emph{entailment} relationship, as the replaced noun and verb can be more or less restrictive compared to the original.
This inherent disparity of sensitivity may worth be taken into consideration when designing future NLI models.

Besides presenting the summary view, the tool also allows the user to quickly narrow down the selection to single example by filtering via the histogram and scatterplot (see Fig.~\ref{fig:summaryView}).
%
Through exploration of many samples with low stabilities (i.e., high sensitivities values in the plots), the domain experts noticed that many highly unstable outliers are from sentence pairs where the prediction is near the decision boundary (see Fig.~\ref{fig:predictStability}(e), the yellow circle corresponds to a \emph{entailment} prediction that is very close to \emph{neutral}).
However, some predictions, such as the one illustrated in Fig.~\ref{fig:predictStability}(d) can alter the prediction quite drastically with minor perturbation (shown in $d_1$, where the word \textbf{heap} is replaced by \textbf{pile} in the hypothesis sentence. In the following section, we will examine what happened inside the model and hypothesis on the cause of failure (see Fig.~\ref{fig:att2pred}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% perturb input, attention
\subsection{Scenario 2: Examine the Decision Making Process}
The predicted label alone provides limited information. Often, domain experts want to know how does the model arrive at a conclusion. And if the prediction is incorrect, wherein the model does the error occur?
Examining the decision making process is not only instrumental in evaluating the model performance but also essential for hypothesizing improvement strategies for future models.
%
In the NLI model, the three stages (encoder, attention, classifier) work in synergy to produce the prediction.
Therefore, making sense of the prediction involves understanding how different parts of the model affect the final prediction.

In the previous section, we have noticed a minor change of the sentence (replace \emph{pile} with \emph{heap}) lead to difference in the final prediction label (Fig.~\ref{fig:predictStability}(d)). Here, we want to make sense of what lead to the failed prediction. In this example, the premise \textbf{P} is ``A very young child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow.'', and the original hypothesis \textbf{H1} is ``A child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow.''. The perturbed hypothesis \textbf{H2} replaces the word \textbf{pile} with \textbf{heap} in \textbf{H1}. This example should be rather straightforward for the model since there are only minor differences between the \textbf{P} and \textbf{H1}/\textbf{H2}.

>>>>

In Fig.~\ref{fig:predictStability}(a)(b)(c), we compare the overall prediction stability (regarding synonymous perturbation) for all correct predictions in the development set (10k examples in total).
%
We observe a drastic difference for the stability for \emph{entailment} predictions compare to the \emph{contradiction} and \emph{neutral} ones.
%
Such a distinction can be partially explained by how \emph{entailment} relationship is defined. The relationship is only valid if the concept in the premise is more specific than the concept in the hypothesis. Therefore, the synonymous perturbation may change the \emph{entailment} relationship, as the replaced \emph{noun} or \emph{verb} can be more or less restrictive compared to the original.
This inherent disparity of sensitivity may worth extra consideration when designing future NLI models.

Besides presenting the summary view, the tool also allows the user to quickly narrow down the selection to a single example by filtering via the histogram and scatterplot (see Fig.~\ref{fig:summaryView}).
%
Through the exploration of many samples with low stabilities (i.e., high sensitivities values in the plots), domain experts notice that many highly unstable outliers are from sentence pairs where the predictions are near the decision boundary (see Fig.~\ref{fig:predictStability}(e), the yellow circle corresponds to a \emph{entailment} prediction that is very close to \emph{neutral}).
However, some predictions, such as the one illustrated in Fig.~\ref{fig:predictStability}(d) can alter the prediction quite drastically with minor perturbation (shown in $d_1$, where the word \textbf{heap} is replaced by \textbf{pile} in the hypothesis sentence. In the following section, we will examine what happened inside the model and hypothesis on the cause of failure (see Fig.~\ref{fig:att2pred}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% perturb input, attention
\subsection{Scenario 2: Examine the Decision Making Process}
The predicted label alone provides limited information. Often, domain experts want to know how does the model arrive at a conclusion. And if the prediction is incorrect, where in the model does the error occur?
Examining the decision making process is not only instrumental in evaluating the model performance but also essential for hypothesizing improvement strategies for future models.
%
In the NLI model, the three stages (encoder, attention, classifier) work in synergy to produce the prediction.
Therefore, making sense of the prediction involves understanding how different parts of the model affect the final prediction.

In the previous section, we have noticed that a minor perturbation of the sentence (replace \emph{pile} with \emph{heap}) may result in change to the final prediction (Fig.~\ref{fig:predictStability}(d)). Here, we want to make sense of what lead to the failed prediction. In this example, the premise \textbf{P} is ``A very young child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow.'', and the original hypothesis \textbf{H1} is ``A child in a red plaid coat and pink winter hat makes a snowball in a large \textbf{pile} of snow.''. The perturbed hypothesis \textbf{H2} replaces the word \textbf{pile} with \textbf{heap} in \textbf{H1}. This example should be rather straightforward for the model since there are only minor differences between the \textbf{P} and \textbf{H1}/\textbf{H2}.
